---
title: "U19_P546_1_linear_modeling.Rmd"
output: html_document
date: "2024-03-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# U19-P546-1
## Ryan Murphy, Dane Grosvenor, Kim Dill-McFarland, Matt Altman

## Include Viral Counts and Metadata

### Load libraries

```{r load_libs, message=FALSE}

library(tidyverse)
library(BIGpicture)
library(kimma)
library(RNAetc)
library(SEARchways)
library(ggrepel)
library(limma)
library(edgeR)
library(readr)
library(stringr)
library(lme4)
library(data.table)
library(patchwork)
library(scales)
library(WGCNA)
library(rrvgo)
library(readxl)
library(openxlsx)
library(emmeans)
library(ggupset)
library(ggpubr)

set.seed(101)

```

### Functions

```{r fxs}

loadRData <- function(fileName) {
  load(fileName)
  get(ls()[ls() !="fileName"])
}

```

### Make Directories

```{r make_dirs}

dir.create("dat_clean/linear_model_output")

```

### Load Data

```{r load_dat, message=FALSE}

# load prior voom object, focusing on Mast Cell
mc_voom_viral <- loadRData(file = "dat_clean/voom_objects/mc_voom_viral.RData")
ens_gene_dict <- loadRData(file = "dat_clean/ensembl_gene_dictionary.RData")

```

## 2 Separate Models

## ~ Timepoint*Treatment + median_cv_coverage

```{r time_treat_lm, warning = FALSE, message = FALSE}

mc_viral_lm_timepoint_treatment <- kimma::kmFit(mc_voom_viral,
             model = "~ Timepoint*Treatment + median_cv_coverage",
             run_contrast = TRUE,
             contrast_var = "Timepoint:Treatment",
             run_lm = TRUE,
             metrics = TRUE,
             use_weights = TRUE)

# View outpust of significant DEGs for contrats of interest at range of FDR cutoffs
# summarize_kmFit(mc_viral_lm_timepoint_treatment$lm.contrast, fdr_cutoff = c(0.1, 0.2, 0.25, 0.3)) %>% 
#   filter(contrast_ref == "Baseline Alone" & contrast_lvl %in% c("4 hrs Alone", "8 hrs Alone", "28 hrs Alone",
#                                                                 "52 hrs Alone", "4 hrs AECs", "8 hrs AECs",
#                                                                 "28 hrs AECs", "52 hrs AECs") |
#   contrast_ref == "8 hrs Alone" & contrast_lvl == "8 hrs AECs" |
#   contrast_ref == "28 hrs Alone" & contrast_lvl == "28 hrs AECs" |
#   contrast_ref == "52 hrs Alone" & contrast_lvl == "52 hrs AECs")

# save linear model output
# 06/20/24
saveRDS(mc_viral_lm_timepoint_treatment, file = "dat_clean/linear_model_output/mc_viral_lm_timepoint_treatment.RDS")

```

### ~ viral (log2CPM) + median_cv_coverage

```{r viral_load_lm, warning = FALSE, message = FALSE}

mc_viral_lm_viralloadCPMLog2 <- kimma::kmFit(mc_voom_viral,
                                       model = "~ normSalmonQuantCountsCPMLog2 + median_cv_coverage",
                                       run_contrast = FALSE,
                                       run_lm = TRUE,
                                       metrics = TRUE,
                                       use_weights = TRUE)

# View outpust of significant DEGs at range of FDR cutoffs
# summarize_kmFit(mc_viral_lm_viralloadCPMLog2$lm, fdr_cutoff = c(0.1, 0.2, 0.25, 0.3))

# save linear model output
# 06/20/24
saveRDS(mc_viral_lm_viralloadCPMLog2, file = "dat_clean/linear_model_output/mc_viral_lm_viralloadCPMLog2.RDS")

```

## Create output data frames of linear model outputs

### Mast Cell linear model output

```{r mc_lm_out_df}

mc_viral_lm_timepoint_treatment_out <- mc_viral_lm_timepoint_treatment$lm.contrast %>% 
  tidyr::drop_na() %>% 
  # filter for contrasts of interest
  dplyr::filter(contrast_ref == "Baseline Alone" & contrast_lvl %in% c("4 hrs Alone", "8 hrs Alone", "28 hrs Alone",
                                                                "52 hrs Alone", "4 hrs AECs", "8 hrs AECs",
                                                                "28 hrs AECs", "52 hrs AECs") |
  contrast_ref == "8 hrs Alone" & contrast_lvl == "8 hrs AECs" |
  contrast_ref == "28 hrs Alone" & contrast_lvl == "28 hrs AECs" |
  contrast_ref == "52 hrs Alone" & contrast_lvl == "52 hrs AECs") %>% 
  # create contrasts variable for each pairwise contrast of interest, set as factor
  dplyr::mutate(contrasts = paste0(contrast_ref, "+", contrast_lvl)) %>% 
  dplyr::mutate(contrasts = factor(contrasts, levels = c("Baseline Alone+4 hrs Alone",
                                                  "Baseline Alone+8 hrs Alone",
                                                  "Baseline Alone+28 hrs Alone",
                                                  "Baseline Alone+52 hrs Alone",
                                                  "Baseline Alone+4 hrs AECs",
                                                  "Baseline Alone+8 hrs AECs",
                                                  "Baseline Alone+28 hrs AECs",
                                                  "Baseline Alone+52 hrs AECs",
                                                  "8 hrs Alone+8 hrs AECs",
                                                  "28 hrs Alone+28 hrs AECs",
                                                  "52 hrs Alone+52 hrs AECs"))) %>% 
  dplyr::filter(FDR < 0.25) %>% 
  dplyr::group_by(contrasts) %>% 
  dplyr::arrange(FDR, .by_group = TRUE) %>% 
  dplyr::ungroup()

# save as csv
# 06/20/24
write.csv(mc_viral_lm_timepoint_treatment_out, file = "dat_clean/linear_model_output/mc_viral_lm_timepoint_treatment_out_df.csv")

```

### Viral Load linear model output

```{r viral_load_lm_out_df}

viral_lm_viral_loadCPMLog2_out <- mc_viral_lm_viralloadCPMLog2$lm %>% 
  dplyr::filter(variable == "normSalmonQuantCountsCPMLog2") %>% # filter to just effect of viral load
  dplyr::filter(FDR < 0.25)

# save as csv
# 06/20/24
write.csv(viral_lm_viral_loadCPMLog2_out, file = "dat_clean/linear_model_output/viral_lm_viral_loadCPMLog2_out_df.csv")

```

### Get gene of interest contrasts
IL3RA, Il5RA, IL18R1, IL1RL1, CCR2, CCR5, LTC4S

```{r}

goi_contrasts <- mc_viral_lm_timepoint_treatment$lm.contrast %>% 
  rename(ensembl_gene_id = gene) %>% 
  left_join(ens_gene_dict, by = "ensembl_gene_id") %>% 
  filter(geneName %in% c("IL3RA", "IL5RA", "IL18R1", "IL1RL1", "CCR2", "CCR5", "LTC4S")) %>% 
  filter(FDR < 0.30) %>% 
  relocate(geneName, .after = "ensembl_gene_id") %>% 
  dplyr::filter(contrast_ref == "Baseline Alone" & contrast_lvl %in% c("4 hrs Alone", "8 hrs Alone", "28 hrs Alone",
                                                                "52 hrs Alone", "4 hrs AECs", "8 hrs AECs",
                                                                "28 hrs AECs", "52 hrs AECs") |
  contrast_ref == "8 hrs Alone" & contrast_lvl == "8 hrs AECs" |
  contrast_ref == "28 hrs Alone" & contrast_lvl == "28 hrs AECs" |
  contrast_ref == "52 hrs Alone" & contrast_lvl == "52 hrs AECs") %>% 
  # create contrasts variable for each pairwise contrast of interest, set as factor
  dplyr::mutate(contrasts = paste0(contrast_ref, "+", contrast_lvl)) %>% 
  dplyr::mutate(contrasts = factor(contrasts, levels = c("Baseline Alone+4 hrs Alone",
                                                  "Baseline Alone+8 hrs Alone",
                                                  "Baseline Alone+28 hrs Alone",
                                                  "Baseline Alone+52 hrs Alone",
                                                  "Baseline Alone+4 hrs AECs",
                                                  "Baseline Alone+8 hrs AECs",
                                                  "Baseline Alone+28 hrs AECs",
                                                  "Baseline Alone+52 hrs AECs",
                                                  "8 hrs Alone+8 hrs AECs",
                                                  "28 hrs Alone+28 hrs AECs",
                                                  "52 hrs Alone+52 hrs AECs"))) %>% 
  dplyr::arrange(geneName)

# 06/20/24
write.csv(goi_contrasts, file = "dat_clean/goi_contrasts.csv")

```

### R session info

```{r sesh}

sessionInfo()

```

