---
title: "P546_1_clean_to_voom.Rmd"
output: html_document
date: "2024-03-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# U19-P546-1
## Ryan Murphy, Dane Grosvenor, Kim Dill-McFarland, Matt Altman

## Clean Data to Voom Object

### Load libraries

```{r load_libs, message = FALSE}

library(tidyverse)
library(BIGpicture)
library(kimma)
library(RNAetc)
library(SEARchways)
library(ggrepel)
library(limma)
library(edgeR)
library(readr)
library(stringr)
library(lme4)
library(data.table)
library(org.Hs.eg.db)
library(patchwork)
library(scales)

set.seed(101)

```

### Functions

```{r fxs}

# https://stackoverflow.com/a/39671052
colClean <- function(x){ colnames(x) <- gsub(".*lib", "lib", colnames(x)); x } 

```

### Make Directories

```{r make_dirs}

dir.create("dat_clean", showWarnings = FALSE)
dir.create("dat_clean/voom_objects")
dir.create("figures", showWarnings = FALSE)
dir.create("figures/PCA", showWarnings = FALSE)

```

### Load Data

```{r load_dat, warning = FALSE}

p546_counts <- read.csv(file = "dat_raw/230727-P546-1_FilteredNormGeneCounts.csv")
p546_raw_counts <- read.csv(file = "dat_raw/230727-P546-1_RawGeneCounts.csv")
p546_QCMetrics <- read.csv(file = "dat_raw/230727-P546-1qcMetricsSummary.csv")
p546_anno <- read.csv(file = "dat_raw/P546-1_Final_Annotation.csv")

# join QC metadata and annotation metadata
p546_meta <- p546_anno %>%
  dplyr::rename(libid = lib.ID) %>%
  left_join(p546_QCMetrics, by = "libid") %>%
  dplyr::rename(libID = libid)

```

### Ensembl Database

```{r get_ensembl, mesage = FALSE}

# load ensembl key from initial runs 03/26/2024
key <- read.csv(file = "dat_raw/ensembl_key_03_26_2024.csv")

# ensembl <- biomaRt::useEnsembl(biomart="ensembl", dataset="hsapiens_gene_ensembl",
#                                 mirror = "useast")
# 
# key <- biomaRt::getBM(attributes=c("ensembl_gene_id", "entrezgene_id", "hgnc_symbol",
#                                         "gene_biotype", "chromosome_name",
#                                         "start_position", "end_position"), mart=ensembl) %>%
#   #Filter protein coding genes
#   filter(gene_biotype == "protein_coding")

# write.csv(key, "dat_raw/ensembl_key_03_26_2024.csv")

```

### Data Adjustment

```{r dat_adj}

# get column names (libIDs) of counts data, remove ensembl ID and gene symbol
counts_libIDs <- colnames(p546_counts)[-c(1:2)]

# adjust metadata
p546_meta <- p546_meta %>% 
  dplyr::filter(libID != "") %>% 
  dplyr::mutate(lib.ID = libID) %>% 
  tibble::column_to_rownames(var = "lib.ID") %>% 
  dplyr::filter(libID %in% counts_libIDs)

# make gene symbol : ensembl ID dictionary
ens_gene_dict <- p546_raw_counts %>% 
  dplyr::filter(geneBiotype == "protein_coding") %>% 
  dplyr::select(geneId, geneName, geneBiotype) %>% 
  mutate(ensembl_gene_id = geneId)

save(ens_gene_dict, file = "dat_clean/ensembl_gene_dictionary.RData")

# filter raw counts to protein coding
raw_counts_filtered <- p546_raw_counts %>%
  mutate(ensembl_gene_id = geneId) %>%
  filter(geneBiotype == "protein_coding") # 19,177 ensemblIDs

# adjust sample colnames
raw_counts_filtered <- colClean(raw_counts_filtered)

# check duplicate ensembl IDs
dup_ensembls <- data.frame(table(raw_counts_filtered$geneId)) %>%
  dplyr::filter(Freq > 1) %>%
  pull(Var1)

# add symbol suffix to dup ensembl IDs
raw_counts_filtered_dups <- raw_counts_filtered %>%
  mutate(geneId = case_when(geneId %in% dup_ensembls ~ paste0(geneId, "_", geneName),
                            TRUE ~ geneId))

# adjust data frame for edgeR and voom requirements
p546_metaV2 <- p546_meta %>% 
  relocate(libID, .before = `Proj.Sub`) %>%
  relocate(`Donor.ID`, .after = libID)

# remove extra cols and force to matrix w/ ensembl IDs as rownames
p546_counts_raw <- raw_counts_filtered_dups %>%
  dplyr::select(-c(geneName, description, geneBiotype, ensembl_gene_id)) %>%
  column_to_rownames(var = "geneId") %>%
  as.matrix()

# sanity check
identical(colnames(p546_counts_raw), rownames(p546_metaV2)) # TRUE

```

### PCA (AECs and MCs)

```{r pca_all, message = FALSE}

pca_all_cells_before_norm <- BIGpicture::plot_pca(
  p546_counts_raw,
  meta = p546_metaV2,
  vars = c("Cell.Type", "Treatment", "Stimulation", "Timepoint", 
           "median_cv_coverage", "outlier")
)

pca_all_cells_before_norm <- pca_all_cells_before_norm$Cell.Type + pca_all_cells_before_norm$Treatment +
  pca_all_cells_before_norm$Stimulation + pca_all_cells_before_norm$Timepoint +
  pca_all_cells_before_norm$median_cv_coverage + pca_all_cells_before_norm$outlier

ggsave(pca_all_cells_before_norm, file = "figures/PCA/PCA_all_cells_before_norm.png") 

```

### Filter poor quality libraries (AECs and MCs)

```{r lib_qc}

seq_cutoff <- 1e6
cv_cutoff <- 1
align_cutoff <- 75

# all libraries pass seq cutoff
ggplot(p546_metaV2,
       aes(x = reorder(libID, fastq_total_reads), y = fastq_total_reads)) +
  geom_col() +
  geom_hline(yintercept = seq_cutoff) +
  scale_y_continuous(trans = 'log10',
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x))) +
  #Beautify
  theme_classic() +
  labs(x = "Library", y = "Pass-filter sequences (log scale)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#Set x-axis max to the larger of 1 or the highest data value
CV_max <- max(1, max(p546_metaV2$median_cv_coverage))

#Set cutoffs to label with libID
cv_cutoff_label <- 0.9
align_cutoff_label <- 75

ggplot(p546_metaV2, aes(
  x = median_cv_coverage, 
  y = mapped_reads_pct)) +
  geom_point() + 
  #Rescale axis limits
  lims(x = c(0,CV_max), y=c(0,100)) +
  #Add cutoff lines
  geom_hline(yintercept = align_cutoff, lty="dashed") +
  geom_vline(xintercept = cv_cutoff, lty="dashed") +
  #Label points outside cutoffs
  geom_text_repel(data=filter(p546_metaV2, median_cv_coverage > cv_cutoff_label | 
                              mapped_reads_pct < align_cutoff_label),
            aes(label=libID), show.legend = FALSE, max.overlaps = Inf,
            box.padding = 1) +
  #Beautify
  theme_classic() +
  labs(x = "Median CV coverage", y="Percent alignment")

```

## Split Data by Cell Type and Run Voom Transformation
### AEC

```{r aec_dat}

# filter to AEC cell type
aec_meta <- p546_metaV2 %>%
  dplyr::filter(`Cell.Type` == "Airway Epithelial Cells")

# use "p546_counts_raw" to subset AEC
# get AEC lib IDs
aec_libs <- rownames(aec_meta)

# filter raw counts to only AECs
aec_counts_raw <- p546_counts_raw %>% 
  as.data.frame() %>% 
  dplyr::select(all_of(aec_libs))

# sanity check
identical(colnames(aec_counts_raw), rownames(aec_meta)) # TRUE

```

### Rare Gene Filter and Voom Transformation

```{r aec_voom}

# make vars of interest as factors, level factors
aec_meta <- aec_meta %>% 
  mutate(Timepoint = factor(Timepoint, levels = c("Baseline", "8 hrs", "28 hrs", "52 hrs"))) %>%
  mutate(Treatment = as.factor(Treatment)) %>%
  mutate(Stimulation = factor(Stimulation, levels = c("None", "HRV-Infected MCs")))

# dge list
aec_dge_list <- edgeR::DGEList(
  counts = aec_counts_raw,
  samples = aec_meta,
  genes = ens_gene_dict
  )

# filter low abundance genes
# min sample ~10% (3)
aec_dge_list_filter <- RNAetc::filter_rare(aec_dge_list, min.CPM = 0.50, min.sample = 3,
                    gene.var = "geneName")

# calculate scaling factors
calc_norm_factors_aec <- edgeR::calcNormFactors(aec_dge_list_filter)

# voom transformation
aec_voom <- limma::voom(
  counts = calc_norm_factors_aec,
  plot = TRUE
)

# save object
# 06/20/2024
save(aec_voom, file = "dat_clean/voom_objects/aec_voom.RData")

```

### Mast Cell

```{r mc_dat}

# filter to only MCs
mc_meta <- p546_metaV2 %>%
  filter(`Cell.Type` == "LAD2 MCs")

# get MC lib IDs
mc_libs <- rownames(mc_meta)

# filter raw counts to only MCs
mc_counts_raw <- p546_counts_raw %>% 
  as.data.frame() %>% 
  dplyr::select(all_of(mc_libs))

# sanity check
identical(colnames(mc_counts_raw), rownames(mc_meta)) # TRUE

```

### Rare Gene Filter and Voom Transformation

```{r mc_voom}

# filter protein coding genes
mc_counts_raw_filtered <- mc_counts_raw %>%
  filter(rownames(.) %in% key$ensembl_gene_id)

# mutate vars of interest to factors and factor levels
mc_meta <- mc_meta %>%
  mutate(Timepoint = factor(Timepoint, levels = c("Baseline", "4 hrs", "8 hrs", "28 hrs", "52 hrs"))) %>%
  mutate(Treatment = factor(Treatment, levels = c("Alone", "AECs"))) %>%
  mutate(Stimulation = factor(Stimulation, levels = c("None", "HRV Infection")))

# dge list
mc_dge_list <- edgeR::DGEList(
  counts = mc_counts_raw_filtered,
  samples = mc_meta
)

# Filter low abundance genes
# min sample ~10% (3)
mc_dge_list <- RNAetc::filter_rare(mc_dge_list, min.CPM = 0.1, min.sample = 3,
                    gene.var = "symbol")

calc_norm_factors_mc <- edgeR::calcNormFactors(mc_dge_list)

mc_voom <- limma::voom(
  counts = calc_norm_factors_mc,
  plot = TRUE
)

# uncomment if need for re-run, keep current version 05/23/24

# re-run 06/20/24
save(mc_voom, file = "dat_clean/voom_objects/mc_voom.RData")

```

### R session info

```{r sesh}

sessionInfo()

```








